---
title: "Plots_Master"
author: "Aji John and Kavya Pradhan"
date: '2024-08-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This code is to create all the plots that we will use in publications/presentations


```{r workspace, message = F, echo = F}
rm(list=ls())
#First load all necessary packages
library(ggplot2)
library(tidyverse)
library(lme4)
library(sjPlot)
library(patchwork)
library(ggpubr)
```

### Raw temperature plots
#### 1. Load data for raw plots
```{r load offset data, message = F, warning = F}
indir <- getwd()
regional_Trim <- read.csv(paste0(indir, "/data/plottingData/regional_Trim.csv"))
local_Trim <- read.csv(paste0(indir, "/data/plottingData/local_Trim.csv"))
localWithHobo <- read.csv(paste0(indir, "/data/plottingData/localWithHobo.csv"))
regionalWithHobo <- read.csv(paste0(indir, "/data/plottingData/regionalWithHobo.csv"))
mmet_june <- read_csv("./data/plottingData/mmetjune2021.csv")
```

#### 2. Fix factors for ploting
```{r plot raw temps, message = F, warning = F}
localWithHobo$TRT <- factor(localWithHobo$TRT, levels = c("OG", "CON", "RClearing", "THIN"))
regional_Trim_augmented <- regional_Trim %>% mutate(Day= lubridate::day(Date)) %>% as.data.frame()
regionalWithHobo$TRT <- factor(regionalWithHobo$TRT, levels = c("OG", "CON", "RClearing", "THIN"))

```

### Comparative Plots

```{r, tidy=FALSE,warning=FALSE,error=FALSE,error=FALSE}

MTControlThinOG<- localWithHobo %>%   filter(Day %in%  c(23:30)) %>%  
  group_by(Day,TRT) %>%
   summarise(mt = mean(Temp)) %>% 
   dplyr::select(TRT,Day,mt) %>% as.data.frame()


colnames(MTControlThinOG) <- c("TRT" ,"Day",       "mt")

MTRegional   <- regional_Trim_augmented %>% mutate(TRT='Regional',Day = lubridate::day(Date),mth= lubridate::month(Date)) %>% filter(Day %in%  c(23:30), mth== 6) %>% 
  dplyr::select(TRT,Day,tmeanC) %>% as.data.frame()
colnames(MTRegional) <- c("TRT" ,"Day",       "mt")

MTClearing   <- mmet_june %>% mutate(TRT='Clearing') %>% filter(day %in%  c(23:30)) %>% group_by(day,TRT) %>%
  summarise(mt = mean(MeanMM))  %>% 
  dplyr::select(TRT,day,mt) %>% 
  as.data.frame()

colnames(MTClearing) <- c("TRT" ,"Day",       "mt")

MT <- rbind(MTControlThinOG,MTClearing,MTRegional)
posterPal5 <- c("#414487FF", "#2A788EFF", "#7AD151FF", "#ffc107",  "#FDE725FF")
posterPal5 <- c("#006d2c", "#41b6c4", "#7AD151FF", "#ffc107",  "#FDE725FF")
posterPal5 <- c("#238443", "#004529", "#ffc107", "#7AD151FF", "#FDE725FF")
posterPal5 <- c("#ffc107", "#238443", "#004529", "#FDE725FF", "#7AD151FF")
MT %>% filter(Day %in% c(23:30)) %>%
  ggplot()+
  annotate("rect", xmin=24,
           xmax = 28,
           ymin = -Inf,
           ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_smooth(aes(Day,mt,color=TRT, linetype = TRT),se=F, size = 3) +
  geom_point(aes(Day,mt,color=TRT, shape = TRT, fill = TRT), size = 7) +
  
  ggpubr::theme_pubclean(base_size = 50) +
  scale_colour_manual(
    values = posterPal5,
    labels = c("Clearing"="Clear-cut", "CON"="Control", "Regional","OG"="Old growth","THIN"="Thinned")
  )  +
  labs(x= "Time (Day in June)", y= "Daily Mean Temperature (°C)",color="") + 
  scale_shape_manual(values = c(15,16,17,23,25), guide = 'none') + scale_fill_manual(values = posterPal5, guide = 'none') + scale_x_continuous(breaks = seq(23, 30, by =1)) + scale_linetype_manual(values = c(1,2,3,4,5), guide = 'none')

   

ggsave("figs/fig3b-alltemps-comparemeans.jpg", width = 45, height = 32, units = "cm")
ggsave("figs/fig3b-alltemps-comparemeans.pdf", width = 45, height = 32, units = "cm")
``` 


### Hottest day map

#### 1. Load the data for mapping
```{r, message = F, warning = F}
library(AOI)
library(climateR)
# 1. Open jpeg file
png(filename = "figs/wahotday.png")
AOIWA = aoi_get(state = "WA")
#AOIWA = aoi_get(list(47,-124 , 200,300)) # lat, long, height, width # change this to change extent of grid data
TempHot = AOIWA %>% getGridMET(varname = "tmmx", startDate = "2021-06-27")
#> Spherical geometry (s2) switched off
#> Spherical geometry (s2) switched on
TempHot$gridmet_tmax_C <- weathermetrics::kelvin.to.celsius(TempHot$daily_maximum_temperature)

aoi_ells <- raster::extent(c(-123.92280 ,-123.88176,46.35758  , 46.42449))
p <- as(aoi_ells, 'SpatialPolygons') 
```

#### 2. Plot the hottest day with ggplot
```{r, message = F, warning = F}
library(terra)
library(tidyterra)
ggplot() +
  geom_spatraster(data = TempHot$gridmet_tmax_C, aes(fill = `tmmx_2021-06-27`)) +
  # You can use coord_sf
  coord_sf(crs = 4326) +
  scale_fill_hypso_c() +
  ggpubr::theme_pubclean()+
  labs(fill="Tmax C")
ggsave("figs/fig3a-hottestDay_ggplot.pdf", width = 30, height = 32, units = "cm")
```

#### 2. Make the hottest day map with raster is
```{r, message = F, warning = F}
hottestDay <- rasterVis::levelplot(TempHot$daily_maximum_temperature, margin = FALSE, main = "Washington State, 06/27/2021", col.regions = viridisLite::viridis(100)) +
  latticeExtra::layer(sp::sp.polygons(p), packets = 1)
hottestDay
ggsave("figs/3a-alt-hottestDay.pdf", width = 45, height = 32, units = "cm")
```

### Offset plots
#### 1. Load data for offset plots

```{r load offset data, message = F, warning = F}
indir <- getwd()
localWithHobo <- read.csv(paste0(indir, "/data/plottingData/localWithHobo.csv"))
regionalWithHobo <- read.csv(paste0(indir, "/data/plottingData/regionalWithHobo.csv"))
```


#### 3. Regional plot
Plot the raw data for regional comparision
Plot the raw temperatures
```{r plot raw temps, message = F, warning = F}

regionalWithHobo$TRT <- factor(regionalWithHobo$TRT, levels = c("OG", "CON", "RClearing", "THIN"))

posterPal <- c("#238443", "#004529", "#ffc107", "#7AD151FF")
posterPal <- c("#239000", "#004529", "#ffc107", "#7AD140" )

rawTempsStudy <- regionalWithHobo %>%  group_by(Day,TRT) %>%
  
  summarise(mt = mean(Temp)) %>%
  ggplot() + 
  annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,mt,color=TRT), size = 7) +
  geom_line(aes(Day,mt,color=TRT), size = 2) +
  geom_point(aes(Day,tmeanC,color='Regional'), data=regional_Trim_augmented, size = 7) +
  geom_line(aes(Day,tmeanC,color='Regional'), data=regional_Trim_augmented, size = 2) +
  
  ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= " Mean Temperature (°C)",color="")+
  scale_x_continuous(breaks = seq(1,30, by = 2)) +
  scale_colour_manual(
    values = posterPal,
    labels = c("CON"="Control", "Regional"="Regional", "THIN"="Thinned", "OG" = "Old Growth")
  )

```

#### 2. Local plot
Plot the raw data for local comparision
Plot the raw temperatures
```{r plot raw temps, message = F, warning = F}
posterPal4 <- c("#238443", "#004529", "#ffc107", "#7AD151FF")
posterPal4 <- c("#239000", "#004529", "#ffc107", "#7AD140" )
rawTempsLocalStudy <- localWithHobo %>%  mutate(typ=Type ) %>% group_by(Day,TRT,typ) %>%
  
   summarise(mt = mean(Temp)) %>%
      ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD")  +
  geom_point(aes(Day,mt,color=TRT, shape = typ), size = 7) +
  geom_line(aes(Day,mt,color=TRT,linetype=typ), size = 3, show.legend = F) +
  geom_point(aes(Day,MeanMean,color='RClearing', shape = Type), data=local_Trim, size = 7) +
  geom_line(aes(Day,MeanMean,color='RClearing',linetype=Type), data=local_Trim, size = 3, show.legend = F) +
  
  ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= " Mean Temperature (°C)",color="",linetype="")+
scale_x_continuous(breaks = c(1:30)) +
 scale_colour_manual(
    values = posterPal4,
 labels = c("CON"="Control", "RClearing"="Clear-cut", "THIN"="Thinned", "OG" = "Old Growth")
)  +
  scale_shape_manual(
    values = c(17,19),labels=c("Air"="Air","Soil"="Surface")
) +
  scale_linetype_manual(values = c("solid","dashed"),) + theme(legend.box="vertical", legend.box.margin=margin(-10,-10,-10,-10), legend.margin=margin(0,0,0,0), legend.title = element_blank())

```

#### 2. Local plots

Plot the offsets for Mean. Note that greater offset == more cooling effect
```{r plot mean offset, message = F, warning = F}
posterPal3 <- c("#239000", "#004529", "#7AD140")
meanLocalOffset <- 
  localWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MeanOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT),shape=17, size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT),linetype=1, size = 3, show.legend = F) +
  geom_point(aes(Day,Soil,color=TRT) ,shape=19, size = 7, show.legend = F) +
  geom_line(aes(Day,Soil,color=TRT),linetype=2, size = 3, show.legend = F) + 
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Mean Cooling effect\n (°C)",color="")+
scale_x_continuous(breaks = seq(1:30)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)

```


Plot the offsets for Diel variation 
```{r plot diel offset, message = F, warning = F}
diurnalLocalOffset <- localWithHobo %>% mutate(typ=Type ) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(RangeOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD")+
  geom_point(aes(Day,Air,color=TRT),shape=17, size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT),linetype=1, size = 3 , show.legend = F) +
  geom_point(aes(Day,Soil,color=TRT) ,shape=19, size = 7 , show.legend = F) +
  geom_line(aes(Day,Soil,color=TRT),linetype=2, size = 3 , show.legend = F) + 

 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Offset Diel Variation \n (°C)",color="")+
scale_x_continuous(breaks = c(1:30))  +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
) + scale_linetype_manual(values=c("solid", "longdash"))
```

Plot the offsets for min
```{r plot minn offset, message = F, warning = F}
minLocalOffset <- 
  localWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MinOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT),shape=17, size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT),linetype=1, size = 3 , show.legend = F) +
  geom_point(aes(Day,Soil,color=TRT) ,shape=19, size = 7 , show.legend = F) +
  geom_line(aes(Day,Soil,color=TRT),linetype=2, size = 3 , show.legend = F) + 
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Min Cooling effect \n (°C)",color="")+
scale_x_continuous(breaks = c(1:30)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)

```

Plot the offsets for max
```{r plot max offset, message = F, warning = F}
maxLocalOffset <- 
  localWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MaxOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT),shape=17, size = 7  , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT),linetype=1, size = 3 , show.legend = F) +
  geom_point(aes(Day,Soil,color=TRT) ,shape=19, size = 7 , show.legend = F) +
  geom_line(aes(Day,Soil,color=TRT),linetype=2, size = 3 , show.legend = F) + 
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Max Cooling effect \n (°C)",color="")+
scale_x_continuous(breaks = c(1:30)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)

```


#### Regional plots

Plot the offsets for Mean. Note that greater offset == more cooling effect
```{r plot mean offset, message = F, warning = F}
meanRegionalOffset <- 
  regionalWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MeanOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT), size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT), size = 3, show.legend = F) +
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Mean Cooling effect\n (°C)",color="")+
scale_x_continuous(breaks = seq(1,30,2)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)

```

Plots for min
```{r plot mean offset, message = F, warning = F}
minRegionalOffset <- 
  #filter(Day %in% c(14:30)) %>% 
  regionalWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MinOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT), size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT), size = 3, show.legend = F) +
  #geom_point(aes(Day,Soil,color=TRT) ,shape=19, size = 7) +
  #geom_line(aes(Day,Soil,color=TRT),linetype=2, size = 3) + 
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Min Cooling effect\n (°C)",color="")+
scale_x_continuous(breaks = seq(1,30,2)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)
minRegionalOffset

```


Plots for max
```{r plot mean offset, message = F, warning = F}
maxRegionalOffset <- 
  regionalWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(MaxOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT), size = 7 , show.legend = F) +
  geom_line(aes(Day,Air,color=TRT), size = 3, show.legend = F) +
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Max Cooling effect\n (°C)",color="")+
scale_x_continuous(breaks = seq(1,30,2)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)
maxRegionalOffset

```


Plots for diel
```{r plot diurnal regional, message = F, warning = F}
diurnalRegionalOffset <- 
  regionalWithHobo %>%  mutate(typ=Type) %>% group_by(Day,TRT,typ) %>%
  summarise(mtHobs = mean(RangeOffset)) %>% spread(typ,mtHobs) %>%
    ggplot() +
   annotate("rect", xmin=24,
               xmax = 28,
               ymin = -Inf,
               ymax = Inf, alpha=0.5, fill="#BDBDBD") +
  geom_point(aes(Day,Air,color=TRT), size = 7, show.legend = F ) +
  geom_line(aes(Day,Air,color=TRT), size = 3, show.legend = F) +
 ggpubr::theme_pubclean( base_size = 50)+
  labs(x= "Time (Day in June)", y= "Diel variation offset\n (°C)",color="")+
scale_x_continuous(breaks = seq(1,30,2)) +
    scale_colour_manual(
    values = posterPal3,
  labels = c("CON"="Control", "THIN"="Thinned", "OG" = "Old Growth")
)
diurnalRegionalOffset


```



### Raw and Offset together plots
#### 2. Local plot

Then put them together and save
```{r, message = F, warning = F}
(rawTempsLocalStudy|(minLocalOffset / maxLocalOffset / meanLocalOffset / diurnalLocalOffset + plot_layout(guides = 'auto')) )+ plot_layout(guides = 'auto')+
  plot_annotation(tag_levels = 'A')

ggsave("figs/fig4-allLocalPlots.pdf", width = 90, height = 80, units = "cm")
```


####3. Regional plot
Then put them together and save
```{r, message = F, warning = F}
(rawTempsStudy|(minRegionalOffset / maxRegionalOffset / meanRegionalOffset / diurnalRegionalOffset + plot_layout(guides = 'auto')) )+ plot_layout(guides = 'auto')+
  plot_annotation(tag_levels = 'A')

ggsave("figs/fig5-allRegionallPlots.pdf", width = 90, height = 80, units = "cm")
```





